% Copyright (C) 2014-2020 by Thomas Auzinger <thomas@auzinger.name>

\documentclass[draft,final]{vutinfth} % Remove option 'final' to obtain debug information.

% Load packages to allow in- and output of non-ASCII characters.
\usepackage{lmodern}        % Use an extension of the original Computer Modern font to minimize the use of bitmapped letters.
\usepackage[T1]{fontenc}    % Determines font encoding of the output. Font packages have to be included before this line.
\usepackage[utf8]{inputenc} % Determines encoding of the input. All input files have to use UTF8 encoding.

% Extended LaTeX functionality is enables by including packages with \usepackage{...}.
\usepackage{amsmath}    % Extended typesetting of mathematical expression.
\usepackage{amssymb}    % Provides a multitude of mathematical symbols.
\usepackage{mathtools}  % Further extensions of mathematical typesetting.
\usepackage{microtype}  % Small-scale typographic enhancements.
\usepackage[inline]{enumitem} % User control over the layout of lists (itemize, enumerate, description).
\usepackage{multirow}   % Allows table elements to span several rows.
\usepackage{booktabs}   % Improves the typesettings of tables.
\usepackage{subcaption} % Allows the use of subfigures and enables their referencing.
\usepackage[ruled,linesnumbered,algochapter]{algorithm2e} % Enables the writing of pseudo code.
\usepackage[usenames,dvipsnames,table]{xcolor} % Allows the definition and use of colors. This package has to be included before tikz.
\usepackage{nag}       % Issues warnings when best practices in writing LaTeX documents are violated.
\usepackage{todonotes} % Provides tooltip-like todo notes.
\PassOptionsToPackage{hyphens}{url}
\usepackage{hyperref}  % Enables cross linking in the electronic document version. This package has to be included second to last.
%\usepackage[hidelinks]{hyperref} % hidelinks for print version

\hypersetup{bookmarksdepth=3}


%--------% my own settings
\usepackage[sort&compress, capitalise, noabbrev]{cleveref} % options: sort, compress and noabbrev
\usepackage{graphics}
\usepackage{multirow}
\usepackage{tikz}
\usepackage{adjustbox}
\usepackage{float}
\usepackage{graphicx}
\usepackage{rotating}
% algorithms
\usepackage{algorithmic}
\floatname{algorithm}{Procedure}

\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}

\newcommand*\circled[1]{\tikz[baseline=(char.base)]{
    \node[shape=circle,draw,inner sep=1pt] (char) {#1};}}

\usepackage{caption}
\usepackage{subcaption}

\usepackage{makecell} % http://mirrors.ibiblio.org/CTAN/macros/latex/contrib/makecell/makecell.pdf

%\renewcommand{\arraystretch}{1.5}

\newcommand{\blue}[1]{\textcolor{blue}{#1}}
\newcommand{\red}[1]{\textcolor{red}{#1}}

%\usepackage[digitsep={,}]{siunitx} %didnt work
\usepackage{sistyle}
\SIthousandsep{,}

\usepackage{epstopdf}

\usepackage{footnote}
\makesavenoteenv{tabular}
\makesavenoteenv{table}

\usepackage[flushleft]{threeparttable} %for footnotes right after the table

\usepackage{doi}

% change caption fontsize
\captionsetup[figure]{font=small,labelfont=small}
\captionsetup[table]{font=small,labelfont=small}

\DeclareUnicodeCharacter{FB01}{fi}

%--------%

\usepackage[acronym,toc]{glossaries} % Enables the generation of glossaries and lists fo acronyms. This package has to be included last.

% Define convenience functions to use the author name and the thesis title in the PDF document properties.
\newcommand{\authorname}{Isabell Lederer} % The author name without titles.
\newcommand{\thesistitle}{Selection Guidelines for Backdoor-based Model Watermarking} % The title of the thesis. The English version should be used, if it exists.

% Set PDF document properties
\hypersetup{
    pdfpagelayout   = TwoPageRight,           % How the document is shown in PDF viewers (optional).
    linkbordercolor = {Melon},                % The color of the borders of boxes around crosslinks (optional).
    pdfauthor       = {\authorname},          % The author's name in the document properties (optional).
    pdftitle        = {\thesistitle},         % The document's title in the document properties (optional).
    pdfsubject      = {Subject},              % The document's subject in the document properties (optional).
    pdfkeywords     = {a, list, of, keywords} % The document's keywords in the document properties (optional).
}

\setpnumwidth{2.5em}        % Avoid overfull hboxes in the table of contents (see memoir manual).
\setsecnumdepth{subsection} % Enumerate subsections.

\nonzeroparskip             % Create space between paragraphs (optional).
\setlength{\parindent}{0pt} % Remove paragraph identation (optional).

\makeindex      % Use an optional index.
\makeglossaries % Use an optional glossary.
%\glstocfalse   % Remove the glossaries from the table of contents.

% Set persons with 4 arguments:
%  {title before name}{name}{title after name}{gender}
%  where both titles are optional (i.e. can be given as empty brackets {}).
\setauthor{}{\authorname}{BSc}{female}
\setadvisor{Ao.Univ.Prof. Dipl.-Ing. Dr.techn.}{Andreas Rauber}{}{male}

% For bachelor and master theses:
\setfirstassistant{Univ.Lektor Mag.rer.soc.oec. Dipl.-Ing.}{Rudolf Mayer}{}{male}
%\setsecondassistant{Pretitle}{Forename Surname}{Posttitle}{male}
%\setthirdassistant{Pretitle}{Forename Surname}{Posttitle}{male}

% For dissertations:
%\setfirstreviewer{Pretitle}{Forename Surname}{Posttitle}{male}
%\setsecondreviewer{Pretitle}{Forename Surname}{Posttitle}{male}

% For dissertations at the PhD School and optionally for dissertations:
%\setsecondadvisor{Pretitle}{Forename Surname}{Posttitle}{male} % Comment to remove.

% Required data.
\setregnumber{01526148}
\setdate{15}{09}{2021} % Set date with 3 arguments: {day}{month}{year}.
\settitle{\thesistitle}{Selection Guidelines for Backdoor-based Model Watermarking} % Sets English and German version of the title (both can be English or German). If your title contains commas, enclose it with additional curvy brackets (i.e., {{your title}}) or define it as a macro as done with \thesistitle.
\setsubtitle{}{} % Sets English and German version of the subtitle (both can be English or German).

% CHOOSE TITLE -----------------------------------------
% 1) Watermarking Machine Learning models: Guidelines on how to choose a fitting method
% 2) Watermarking Machine Learning models: Selection Guidelines for Backdoor-based Model Watermarking
% 3) Selection Guidelines for Backdoor-based Model Watermarking
% -------------------------------------------------------

% Select the thesis type: bachelor / master / doctor / phd-school.
% Bachelor:
%\setthesis{bachelor}
%
% Master:
\setthesis{master}
\setmasterdegree{dipl.} % dipl. / rer.nat. / rer.soc.oec. / master
%
% Doctor:
%\setthesis{doctor}
%\setdoctordegree{rer.soc.oec.}% rer.nat. / techn. / rer.soc.oec.
%
% Doctor at the PhD School
%\setthesis{phd-school} % Deactivate non-English title pages (see below)

% For bachelor and master:
\setcurriculum{Technical Mathematics}{Technische Mathematik} % Sets the English and German name of the curriculum.

% For dissertations at the PhD School:
\setfirstreviewerdata{Affiliation, Country}
\setsecondreviewerdata{Affiliation, Country}

\begin{document}

\frontmatter % Switches to roman numbering.
% The structure of the thesis has to conform to the guidelines at
%  https://informatics.tuwien.ac.at/study-services

\addtitlepage{naustrian} % German title page (not for dissertations at the PhD School).
\addtitlepage{english} % English title page.
\addstatementpage

%\begin{danksagung*}
%\end{danksagung*}

\begin{acknowledgements*}
In the process of planning, writing and elaborating this thesis, I received broad support from a multitude of people to whom I would like to express my gratitude and acknowledgement.

It goes without saying that I am exceptionally thankful for the guidance by Rudolf Mayer. His advice on methodical approach and research were of fundamental importance, and I highly appreciate his consistent and constructive feedback as well as his collegial ways. Equally, I would like to express my sincere appreciation for the supervision by professor Andreas Rauber and his essential consultation throughout the entire progression of this thesis.

I gratefully address my dear fellow students, especially Karine and Nemanja, my colleagues at SBA Research, and those at the TUtheTOP alumni club for their cordial companionship, motivational support and honest feedback. Many thanks to Tanja and Monika for proofreading this thesis.

I would like to thank my friends and family for the continuous motivation, the interest in the subject area, which is foreign to many of them, and the constant participation, especially on particularly challenging days. I am grateful for my brother’s enthusiasm for what I do and his sincere admiration in seeing me evolve. I thank Alina for reminding me that life is not all about work and taking me out for various activities.

A very special thank you goes, of course, to my partner Florian, who not only endures my ups and downs but also supports me in all my projects, regardless of how big and ambitious they are. I am enormously thankful for his always encouraging words and his great cooking.

Above all, I would like to thank my mother, whose fundamentally positive, unwavering view of the future has given me the will, self-confidence and perseverance without which I would undoubtedly not have been able to cope with the numerous and occasionally hopeless seeming tasks and challenges of my educational path to date, as well as of my private and professional life.
\end{acknowledgements*}

\begin{kurzfassung}
Da die kommerzielle Nutzung von maschinellem Lernen (ML) immer weiter verbreitet ist und die steigende Komplexität von ML-Modellen aufwendiger und damit teurer zu trainieren wird, wächst auch die Dringlichkeit, geistiges Eigentums in diesen Modellen zu schützen. Im Vergleich zu Technologien, die sich auf ein solides Verständnis von Bedrohungen, Angriffen und Verteidigungsmöglichkeiten zum Schutz ihres geistigen Eigentums stützen können, ist die Forschung in dieser Hinsicht bei ML noch sehr fragmentiert. Dies ist mitunter auf das Fehlen einer einheitlichen Sichtweise und einer gemeinsamen Taxonomie dieser Aspekte zurückzuführen.

In dieser Arbeit werden die Erkenntnisse zum Schutz des geistigen Eigentums in ML systematisiert, wobei der Schwerpunkt auf Bedrohungen und Angriffen liegt, die für einige der bisher bestehenden Systeme festgestellt wurden, sowie auf den bisher vorgeschlagenen Schutzmaßnahmen. Wir entwickeln ein umfassendes Bedrohungsmodell für das geistige Eigentum in ML und kategorisieren Angriffe und Abwehrmaßnahmen in einer einheitlichen und konzisen Taxonomie, um auf diese Weise die Brücke zwischen ML und zukunftsweisender Sicherheit zu schlagen.

Später konzentrieren wir uns auf Backdoor-basiertes Watermarking für Deep Neural Networks zur Bildklassifizierung und definieren verschiedene Parameter für eine umfassende Studie dieser Ansätze. Dies ist von grundlegender Bedeutung für die Bewertung der verschiedenen Methoden und die Formulierung des Leitfadens. Schließlich wählen wir eine Teilmenge dieser Parameter aus und vergleichen die Methoden, um eine Empfehlung für eine Watermarking-Methode auf Basis eines ML-Settings zu geben.
\end{kurzfassung}

\begin{abstract}
With commercial uses of Machine Learning (ML) becoming more wide-spread, while at the same time ML models becoming more complex and expensive to train, the Intellectual Property Protection (IPP) of trained models is becoming a pressing issue. Unlike other domains that can build on a solid understanding of the threats, attacks and defences available to protect their IP, the research in this regard in ML is still very fragmented. This is also due to a lack of a unified view and a common taxonomy of these aspects.

In this thesis, we systematise findings on IPP in ML, focusing on threats and attacks identified on these systems and defences proposed to date. We develop a comprehensive threat model for IP in ML, and categorise attacks and defences within a unified and consolidated taxonomy, thus bridging research from both the ML and security communities.

Later on, we focus on backdoor-based watermarking approaches for Deep Neural Networks for image classification and define different parameters and settings for a comprehensive study of these approaches. This will be fundamental for evaluating the different methods and formulating the selection guidelines. Finally, we choose a subset of these parameters and compare the methods in order to provide a recommendation for a watermarking method based on the ML setting.
\end{abstract}

% Select the language of the thesis, e.g., english or naustrian.
\selectlanguage{english}

% Add a table of contents (toc).
\tableofcontents % Starred version, i.e., \tableofcontents*, removes the self-entry.

% Switch to arabic numbering and start the enumeration of chapters in the table of content.
\mainmatter

%Ideen:
% - Problem definition a subsection somewhere.
% - Mache bei den subsections, vorallem black-box watermarking zuerst eine allgemeine übersicht, was es so gibt, und danach erst genauer drauf eingehen und die paper beschreiben...

\include{ch1_introduction}
\include{ch2_methodology}
\include{ch3_background}
\include{ch4_taxonomy}
\include{ch5_stateoftheart}
\include{ch6_studysettings}
\include{ch7_comparison}
\include{ch8_conclusions}
\include{appendix}


\backmatter

% Use an optional list of figures.
\listoffigures % Starred version, i.e., \listoffigures*, removes the toc entry.

% Use an optional list of tables.
\cleardoublepage % Start list of tables on the next empty right hand page.
\listoftables % Starred version, i.e., \listoftables*, removes the toc entry.

% Use an optional list of alogrithms.
%\listofalgorithms
%\addcontentsline{toc}{chapter}{List of Algorithms}

% Add an index.
%\printindex

% Add a glossary.
%\printglossaries

% Add a bibliography.
%\bibliographystyle{ieeetr}
%\bibliographystyle{IEEEtran}
\bibliographystyle{plainurl}
%\bibliographystyle{mybibstyle} % tried but failed -> zotero doesnt export the needed bibtex fields

%\bibliography{bibliography/bibliography.bib, bibliography/model_compression.bib, bibliography/others.bib, bibliography/relatedwork.bib, bibliography/rudi.bib, bibliography/models_and_datasets.bib, bibliography/background.bib}

\bibliography{bibliography/references.bib, bibliography/references2.bib, bibliography/websites.bib}

\end{document}